CuO Analysis Details

### 2019-06-10 paper/draft.tex
finished introduction

\begin{document}
\section{Introduction}

@@

Figure \ref{fig:FinalED} presents the properties and eigenvalues of the lowest 30 eigenstates for three of the six fit models at $\lambda = 20$. 
In comparison one can view Figure \ref{fig:InitED} as the same figure for fits at $\lambda = 0$. 
At $\lambda = 20$ there are no outliers below 2eV in any of the three fit models. 
The ground states are similar for the three models, a feature also seen for the models fit at $\lambda = 0$.
Unlike at $\lambda = 0$, we see that the excited states of the three models are also very similar.
This pattern is also shared by the three other candidate models which are not shown. 
The additional constraint from asserting a prior has seemingly caused a single effective theory to emerge from our set of candidate models.
The values of the fit parameters for our six models can be seen in Table 1.
All the models share similar parameter values except the fifth model, Min$ + \bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi}, \bar{c}_{d_z^2}^\dagger \bar{c}_{4s}$, indicating that there are at least two effective theories on $\mathcal{LE}$ which lead to identical low-energy eigenstates and spectra, describe the energies of our sample data accurately and satisfy our prior.
We believe this arises because our low-energy space has no states which can probe the energy of the excitation Cu 3d $\rightarrow$ O 2p$_z$. 
No single excitation on the ground state can lead to this excitation and the lowest energy double excitation which can probe this difference would be Cu 3d, O 2p$_z \rightarrow$ O 2p$_\pi$, O 4s which would have fewer than nine electrons in the Cu 3d orbitals since since the O 2p$_z$ molecular orbital is hybridized with the Cu 3d$_{z^2}$ orbital.
This energy difference can therefore be freely varied without affecting the quality or eigensolutions to the model, meaning there is no unique effective model for this low-energy space.

@@

\section{Conclusions}
In this paper we systematically fit a many-body model for the neutral CuO molecule which can accurately reproduce the spectrum and properties of low-lying eigenstates seen in experiment using a \textit{ab-initio} QMC DMD method, and discuss the details of and developments in our model fitting procedure.
In particular, we study the problem of intruder states in our effective theories --- states which suffer from large extrapolations errors under our model ---  by connecting the existence of these states to our inability to access parts of $\mathcal{LE}$ when generating the low-energy wave function samples used to fit our models.
The intruder state issue is dealt by using our knowledge of $E_\text{ab}$ on the inaccessible parts of $\mathcal{LE}$ to construct a prior for our regression and enforcing it through an augmented ordinary linear regression cost function.
The resultant models fit using the augmented cost function do not contain any intruder states.
The fit models can accurately describe the energies and electron configurations of experimentally measured low-lying excitations with a much higher accuracy than effective single-particle theories generated by ROKS, for example.
We also discuss the non-uniqueness of effective theories by studying two different effective theories we built which produce almost identical spectra and eigenvectors and fit our sample data accurately.

>>>>

\begin{document}
\section{Introduction}
Systematically developing model Hamiltonians which can accurately describe the energies and properties of the lowest lying eigenstates for strongly correlated electrons systems is a pressing problem in modern condensed matter physics. 
Examples of such materials are high-temperature superconductors and other transition metal oxides.
Low-energy excitations of these systems typically involve collective spin, lattice, and charge degrees of freedom with intermediate couplings.
Consequently, many of these systems have competing exotic long range ordering, for example the hypothesized competition between antiferromagnetic order and d-wave superconducting order in the cuprates.
The main difficulty in describing these materials is that they do not exhibit behavior which can be described by an effective single particle theory nor perturbatively in either weak or strong coupling limits.

The necessity for non-perturbative, many-body models in describing the low-lying excited states of strongly correlated electron systems poses a difficult challenge when developing model Hamiltonians for real world systems.
Fermi liquid theory is the prototypical example of a one-body non-perturbative effective theory for metals.
The low-energy effective theory has long-lived non-interacting single particle excitations near the Fermi surface and the Coulomb interaction is treated non-perturbatively.
The anti-ferromagnetic Heisenberg model is a good example of a many-body perturbative effective theory for Mott insulators.
The low-energy effective theory is built of many-body spin excitations which are derived from perturbing from a strong coupling limit.
These types of models fail when used to model many strongly correlated electron systems, and consequently various non-perturbative many-body models like the t-J, 3-band Hubbard, and Kanamori models have emerged as candidates.
While these models are computational tractable to solve, it is unclear whether they can describe the complexity of real systems, or how one would go about developing a more suitable model for a given material.

Common approaches to building suitable model Hamiltonians for strongly correlated materials involve using single particle theories like density functional theory (DFT).
One-body terms are typically obtained by projecting the solutions from the DFT band structure onto a localized single-particle basis, giving an estimate of hoppings and occupations on a lattice model.
Two-body terms in the model are developed by assuming a screened Coulomb interaction based on constrained DFT or RPA.
Aside from this method, Lowdin methods coupled to a stochastic method and canonical transformations have also been used to develop effective theories.
These methods have the draw back that the constructed effective theory typically does not come with any estimate of the quality of the model, therefore making model validation a phenomenological task which can lead to unclear biases in the model construction.

In this paper we systematically fit a many-body model
which can accurately describe energies and properties of low-lying eigenstates for the neutral CuO molecule using a generalizable \textit{ab-initio} Quantum Monte Carlo (QMC) density matrix downfolding (DMD) method as a stepping stone towards application to bulk strongly correlated materials.
The neutral CuO molecule has been a subject of intense theoretical and experimental study due to the complex structure of its low-energy space.
While the low-energy excitations in the doublet sector of the CuO molecule are well understood, singlet selection rules for spectroscopic measurements and the doublet ground state leave the quartet sector of excitations unexplored.
The CuO molecule is hence a good playground for testing and developing the DMD method as there are some well understood excitations which our model should describe accurately and a set of poorly understood excitations which our model can give some insight into.
The relative simplicity of the system allows us to rigorously detail each step of the DMD procedure, particularly sampling and regression techniques as well as quantifiable model validation.

@@

Figure \ref{fig:FinalED} presents the properties and eigenvalues of the lowest 30 eigenstates for three of the six fit models at $\lambda = 20$. 
In comparison one can view Figure \ref{fig:InitED} as the same figure for fits at $\lambda = 0$. 
At $\lambda = 20$ there are no outliers below 2eV in any of the three fit models. 
The ground states are similar for the three models, a feature also seen for the models fit at $\lambda = 0$.
Unlike at $\lambda = 0$, we see that the excited states of the three models are also very similar.
This pattern is also shared by the three other candidate models which are not shown. 
The additional constraint from asserting a prior has seemingly caused a single effective theory to emerge from our set of candidate models.
The values of the fit parameters for our six models can be seen in Table 1.
All the models share similar parameter values except the fifth model, Min$ + \bar{c}_{d_\pi}^\dagger \bar{c}_{p_\pi}, \bar{c}_{d_z^2}^\dagger \bar{c}_{4s}$, indicating that there are at least two different effective theories on $\mathcal{LE}$ which lead to identical low-energy eigenstates and spectra, describe the energies of our sample data accurately and satisfy our prior.
\textbf{Why does this happen?}

@@

\section{Conclusions}
In this paper we systematically fit a many-body model for the neutral CuO molecule which can accurately reproduce the spectrum and properties of low-lying eigenstates seen in experiment using a \textit{ab-initio} QMC DMD method, and discuss the details of and developments in our model fitting procedure.
In particular, we study the problem of intruder states in our effective theories --- states which suffer from large extrapolations errors under our model ---  by connecting the existence of these states to our inability to access parts of $\mathcal{LE}$ when generating the low-energy wave function samples used to fit our models.
The intruder state issue is dealt by using our knowledge of $E_\text{ab}$ on the inaccessible parts of $\mathcal{LE}$ to construct a prior for our regression and enforcing it through an augmented ordinary linear regression cost function.
The resultant models fit using the augmented cost function do not contain any intruder states.
The fit models can accurately describe the energies and electron configurations of experimentally measured low-lying excitations with a much higher accuracy than effective single-particle theories generated by ROKS, particularly for quartet states.
We also discuss the non-uniqueness of effective theories by studying two different effective theories we built which produce almost identical spectra and eigenvectors and fit our sample data accurately.
